19Q2

### 2019-10-08 19Q2.Rmd

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata, eval=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])

# OLD CCL_META
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")
ccl_meta <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta <- select(ccl_meta, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
ccl_meta <- subset(ccl_meta, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " ", as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == "other" & CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid", disease)))))))

ccl_meta <- merge(merge(ccl_meta, ccl_histologies, by = c("Cell_line", "CCLE_histology", "CCLE_name", "DepMap_ID"), all = TRUE), lineage_converter, by = "DepMap_ID", all = TRUE)
ccl_meta$disease <- with(ccl_meta, ifelse(is.na(disease) & !is.na(CCLE_histology), CCLE_histology, disease))
ccl_meta$disease <- with(ccl_meta, ifelse(CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid",
                                   ifelse(CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(Cell_line == "GT3TKB", "lung",
                                   ifelse(disease == "stomach" | CCLE_histology == "stomach", "gastric", disease))))))))
write.csv(ccl_meta, "./data_munging/19Q2/sample_info_edit_20190718.csv")
```

```{r}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20190717.csv", sep = ",", header = TRUE, row.names = 1)
ccl_meta <- ccl_meta %>% select(DepMap_ID, Cell_line, disease)
ccl_meta <- filter(ccl_meta, DepMap_ID != "ACH-002397")
ccl_unique <- data.frame(disease = unique(ccl_meta$disease))
```

@@

@@

```{r fusion_calls, eval=FALSE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
write.csv.gz(fusions, "./data_munging/19Q2/CCLE_fusions_edit_20190718.csv.gz")
```

```{r}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20190718.csv.gz", header = TRUE, sep = ",")
```

@@

**NEED TO DO THIS**

@@

```{r dg_heatmap, fig.height=6, fig.width=8, warning=FALSE}
g2p_dg_counts <- g2p %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
# g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

hta_annot <- HeatmapAnnotation(text = anno_text(colnames(hta_data), offset = unit(0.1, "in"), gp = gpar(fontsize = 10)))
hta_legend_list <- list(legend_direction = "horizontal", border = "gray70", nrow = 2)
hta <- Heatmap(hta_data,
               name = "Functional screen status           ",
               col = c("Screened" = "purple4", "Not screened" = "gray97"), 
               rect_gp = gpar(col = "gray70"),
               top_annotation = hta_annot,
               heatmap_legend_param = hta_legend_list,
               row_names_side = "left",
               row_names_gp = gpar(fontsize = 10),
               column_names_side = "top",
               column_names_gp = gpar(fontsize = 10),
               show_column_names = FALSE,
               cluster_rows = FALSE, 
               cluster_columns = FALSE,
               row_title = "Drugs",
               row_title_gp = gpar(fontface = "bold", fontsize = 14),
               column_title = "Dataset",
               column_title_gp = gpar(fontface = "bold", fontsize = 14),
               width = unit(1.5, "in"))

@@

@@

```{r auc_clustered_ht, eval=FALSE}
ccle_auc_grid <- dcast(ccle, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
ctrp_auc_grid <- dcast(ctrp, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
gdsc_auc_grid <- dcast(gdsc, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)

@@

>>>>

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccle_meta_test <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
```

```{r depmap_metadata, eval=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])

# OLD CCL_META
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")
ccl_meta_orig <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta_orig_filt <- select(ccl_meta_orig, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta_orig_filt) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
ccl_meta <- subset(ccl_meta_orig_filt, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " ", as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == "other" & CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid", disease)))))))

ccl_meta <- merge(merge(ccl_meta, ccl_histologies, by = c("Cell_line", "CCLE_histology", "CCLE_name", "DepMap_ID"), all = TRUE), lineage_converter, by = "DepMap_ID", all = TRUE)
ccl_meta$disease <- with(ccl_meta, ifelse(is.na(disease) & !is.na(CCLE_histology), CCLE_histology, disease))
ccl_meta$disease <- with(ccl_meta, ifelse(CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid",
                                   ifelse(CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(Cell_line == "GT3TKB", "lung",
                                   ifelse(disease == "stomach" | CCLE_histology == "stomach", "gastric", disease))))))))
# Filter out cell lines without a DepMap ID or a cell line
ccl_meta <- filter(ccl_meta, !is.na(DepMap_ID) & !is.na(Cell_line))
# Filter out cell lines that have a CCLE_name not present in the original sample info from DepMap
ccl_meta <- filter(ccl_meta, CCLE_name %in% ccl_meta_orig$CCLE_name & Cell_line %in% ccl_meta_orig_filt$Cell_line)
# Remove cell line not used in Pozdeyev
ccl_meta <- filter(ccl_meta, DepMap_ID != "ACH-002397")
write.csv(ccl_meta, "./data_munging/19Q2/sample_info_edit_20191008.csv")
```

@@

```{r}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20191008.csv", sep = ",", header = TRUE, row.names = 1)
ccl_meta <- ccl_meta %>% select(DepMap_ID, Cell_line, disease)
ccl_unique <- data.frame(disease = unique(ccl_meta$disease))
```

@@

```{r fusion_calls, eval=FALSE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
write.csv.gz(fusions, "./data_munging/19Q2/CCLE_fusions_edit_20191008.csv.gz")
```

```{r}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20191008.csv.gz", header = TRUE, sep = ",")
```

@@

@@

```{r dg_heatmap, fig.height=6, fig.width=8, warning=FALSE, eval=FALSE}
g2p_dg_counts <- g2p %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
# g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

hta_annot <- HeatmapAnnotation(text = anno_text(colnames(hta_data), offset = unit(0.4, "in"), gp = gpar(fontsize = 10)))
hta_legend_list <- list(legend_direction = "horizontal", border = "gray70", nrow = 2)
hta <- Heatmap(hta_data,
               name = "Functional screen status           ",
               col = c("Screened" = "purple4", "Not screened" = "gray97"), 
               rect_gp = gpar(col = "gray70"),
               top_annotation = hta_annot,
               heatmap_legend_param = hta_legend_list,
               row_names_side = "left",
               row_names_gp = gpar(fontsize = 10),
               column_names_side = "top",
               column_names_gp = gpar(fontsize = 10),
               show_column_names = FALSE,
               cluster_rows = FALSE, 
               cluster_columns = FALSE,
               row_title = "Drugs",
               row_title_gp = gpar(fontface = "bold", fontsize = 14),
               column_title = "Dataset",
               column_title_gp = gpar(fontface = "bold", fontsize = 14),
               width = unit(1.5, "in"))

@@

**ENDED HERE 10/08**

@@

```{r auc_clustered_ht, eval=TRUE}
ccle_auc_grid <- dcast(ccle, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
ctrp_auc_grid <- dcast(ctrp, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
gdsc_auc_grid <- dcast(gdsc, Drug ~ DepMap_ID, value.var = "Score", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)

@@

**10/8 Ran code up to here**
